Article figures

### 2020-11-06 src/article_figures.Rmd
Updated article figures

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
save_figs = FALSE
fig_path="~/FRCBS/results-for-eba-paper/pdf"
```

```{r}

@@

```{r}

@@

```{r}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
male_posterior <- finngenn_male_raw$samples
female_posterior <- finngenn_female_raw$samples
variables <- finngenn_male_raw$pretty_variable_names
# Give a more descriptive name to this variable
variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
```

@@

```{r}
alternative_create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")

@@

  plot <- final %>% 
    ggplot() +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    # ggstance is required to make legend keys horizontal
    geom_rect(data=combined_variables %>% filter(stripe==1),
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                          ymin = as.numeric(Pretty) - 0.5),
              fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA, inherit.aes = FALSE) +
    ggstance::geom_pointrangeh(mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
                               position=position_dodge2(width=0.8, padding=0.1), 
                               size=0.25,  
                               inherit.aes = FALSE) +
    labs(title=NULL, x=NULL, y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female", "both")) + 
    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
    facet_wrap("method", scales="free_x") +
    theme_classic(base_size = base_size) + 
    theme(legend.position = "bottom", legend.direction = "horizontal", strip.background = element_rect(fill = NULL, color = "white", size = 1))

@@

# Show AUROC, AUPR, F1, E6, and E12 as forest plot
create_performance_forest_plot <- function() {
  df <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
  df <- df %>%
    mutate(Id=factor(Id, levels=rev(ids)),
           type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
           sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
           model=factor(str_split_fixed(Id, "-", 3)[,3]))

  thresholds <- get_threshold_df()
  df <- bind_rows(df, thresholds)
  df <- df %>% 
    mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "t6", "t12")))
  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
  xrange <- as.numeric(df %>% filter(type %in% c("E6", "E12")) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
  xrange <- rep(xrange, 2)
  mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, type=mytype)  # Dummy table to use the same x-axis limits for E6 and E12
  #print(dummy)

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model") + 
    #scale_colour_manual(breaks = c("male", "female", "both"), values=c("yellow", "brown", "black"), aesthetics = c("color", "fill", "shape")) +
    scale_colour_discrete(breaks = c("male", "female", "both")) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf"), values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("type", scales="free_x")
  #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
  return(list(g=g, df=df))
}

@@

```{r calibration plots}
calibration_plot <- function(df, use_my_method=FALSE) {
  if (use_my_method) {
    df$cat <- cut(df$Deferred, seq(0.0, 1.0, length.out = 11))
    df2 <- df %>% group_by(cat, .drop=FALSE) %>% summarise(true_bin_prob = mean(obs=="Deferred"))
    df2 %>% ggplot(aes(cat, true_bin_prob)) + geom_point() + ylim(0, 1)
  } else {
    cal <- calibration(obs ~ Deferred, data=df, class="Deferred", cuts=11)
    ggplot(cal) + ylim(0, 100)
  }
}

@@

create_classification_scatter_plot <- function(df, hb_threshold, probability_of_deferral_threshold) {
  xymin <- min(min(df$predicted), min(df$observed))
  xymax <- max(max(df$predicted), max(df$observed))
  df <- df %>% 
    mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0)) %>%
    mutate(confusion_class = factor(ifelse(deferral == 1, 
                                    ifelse(new_predicted_label == 1, "True positive", "False negative"),
                                    ifelse(new_predicted_label == 1, "False positive", "True negative")),
           levels=c("True positive", "False negative", "False positive", "True negative")))
  scatter_plot <- ggplot(df, aes(x = observed, y=predicted, color = confusion_class)) +
    geom_point() +
    #xlim(xymin,xymax) + ylim(xymin,xymax) +
    scale_x_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_y_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    geom_abline(intercept = 0, slope = 1) +
    labs(x = "Observed", y = "Predicted", colour = "Deferral status") +
    #scale_colour_discrete(labels=c("Accepted", "Deferred")) +
    geom_smooth(mapping=aes(x = observed, y=predicted), colour="black", show.legend = FALSE) +
    geom_vline(xintercept = hb_threshold, linetype = "dashed") +
    geom_hline(yintercept = hb_threshold, linetype = "dashed") +
    theme(legend.position = "bottom") +
    ggtitle("Observed vs predicted Hb-values")
  return(scatter_plot)
}
```

```{r}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")

@@

```{r}
g <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.1)
if (save_figs)
  ggsave(filename="classification-scatter-plot-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g, dpi=600, units="mm", width=180)
g
```

```{r, eval=FALSE}
jarkko_test <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
save_figs = TRUE
fig_path="~/FRCBS/results-for-eba-paper/pdf"
```

```{r, include=FALSE}

@@

```{r, include=FALSE}

@@

```{r, include=FALSE}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
male_posterior <- finngenn_male_raw$samples
female_posterior <- finngenn_female_raw$samples
variables <- finngenn_male_raw$pretty_variable_names
# Give a more descriptive name to this variable
variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
```

@@

```{r, include=FALSE}
alternative_create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")

@@

  plot <- final %>% 
    ggplot() +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    # ggstance is required to make legend keys horizontal
    geom_rect(data=combined_variables %>% filter(stripe==1),
              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
                          ymin = as.numeric(Pretty) - 0.5),
              fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA, inherit.aes = FALSE) +
    ggstance::geom_pointrangeh(mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
                               position=position_dodge2(width=0.8, padding=0.1), 
                               size=0.25,  
                               inherit.aes = FALSE) +
    labs(title=NULL, x=NULL, y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) + 
    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
    facet_wrap("method", scales="free_x") +
    theme_classic(base_size = base_size) + 
    theme(legend.position = "bottom", legend.direction = "horizontal", strip.background = element_rect(fill = NULL, color = "white", size = 1))#,
          #plot.background = element_rect(fill = "lightblue"))

@@

get_name_of_data <- function(s) {
  s <- str_split_fixed(s, "-", 3)[,1]
  s <- map(s, function(x) when(x, 
                               x == "progesa" ~ "eProgesa",
                               x == "finngen" ~ "Biobank",
                               x == "findonor" ~ "FinDonor",
                               ~ "koe"))
  return(s)
}

# Show AUROC, AUPR, F1, E6, and E12 as forest plot
create_performance_forest_plot <- function() {
  df <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
  new_names <- c(`Economic effect of 6 month deferral` = "E6", `Economic effect of 12 month deferral` = "E12")
  df <- df %>%
    mutate(Id=factor(Id, levels=rev(ids)),
           type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
           sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
           model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
    mutate(type = fct_recode(type, !!!new_names))

  thresholds <- get_threshold_df()
  df <- bind_rows(df, thresholds)
  df <- df %>% 
    mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "t6", "t12")))
  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
  xrange <- as.numeric(df %>% filter(type %in% c("E6", "E12")) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
  xrange <- rep(xrange, 2)
  mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, type=mytype) # Dummy table to use the same x-axis limits for E6 and E12

  dummy <- dummy %>%
    mutate(type = fct_recode(type, !!!new_names))

  print(df)

  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
  #print(key_size)
  key_size <- 1.8   # This results in box of size 6 mm
  #key_size <- unit(6.096, "mm")
  #key_size <- unit(1.2, "lines")
  #grid::convertX(grid::unit(1.2, "lines"), "mm")
  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model", fill="Deferral length") + 
    scale_y_discrete(labels=get_name_of_data) +
    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:6) + #breaks = new_names) +
    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
                       values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
    #theme(legend.key.size = theme_get()$legend.key.size) +
    geom_blank(data=dummy) +
    facet_wrap("type", scales="free_x") +
    #theme(legend.position = "right", legend.direction = "vertical") 
    theme(legend.position = "bottom", legend.direction = "vertical")
  #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
  return(list(g=g, df=df))
}

@@

```{r calibration plots, include=FALSE}
calibration_plot <- function(df, use_my_method=FALSE) {
  if (use_my_method) {
    df$cat <- cut(df$Deferred, seq(0.0, 1.0, length.out = 11))
    df2 <- df %>% group_by(cat, .drop=FALSE) %>% summarise(true_bin_prob = mean(obs=="Deferred"))
    df2 %>% ggplot(aes(cat, true_bin_prob)) + geom_point() + ylim(0, 1)
  } else {
    cal <- calibration(obs ~ Deferred, data=df, class="Deferred", cuts=11)
    ggplot(cal) + ylim(0, 100)
  }
}

@@

create_classification_scatter_plot <- function(df, hb_threshold, probability_of_deferral_threshold) {
  xymin <- min(min(df$predicted), min(df$observed))
  xymax <- max(max(df$predicted), max(df$observed))
  if (is.null(probability_of_deferral_threshold)) { # predict deferral based on Hb
    df <- df %>% 
      mutate(new_predicted_label= ifelse(predicted < hb_threshold, 1, 0))
  } else { # predict deferral based on probability of deferral
    df <- df %>% 
      mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0))
  }
  df <- df %>%
    mutate(confusion_class = factor(ifelse(deferral == 1, 
                                    ifelse(new_predicted_label == 1, "True positive", "False negative"),
                                    ifelse(new_predicted_label == 1, "False positive", "True negative")),
           levels=c("True positive", "False negative", "False positive", "True negative")))
  scatter_plot <- ggplot(df, aes(x = observed, y=predicted, color = confusion_class)) +
    geom_point() +
    #xlim(xymin,xymax) + ylim(xymin,xymax) +
    scale_x_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_y_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_colour_discrete(drop=FALSE) +
    geom_abline(intercept = 0, slope = 1) +
    labs(x = "Observed", y = "Predicted", colour = "Deferral status") +
    #scale_colour_discrete(labels=c("Accepted", "Deferred")) +
    geom_smooth(mapping=aes(x = observed, y=predicted), colour="black", show.legend = FALSE) +
    geom_vline(xintercept = hb_threshold, linetype = "dashed") +
    geom_hline(yintercept = hb_threshold, linetype = "dashed") +
    theme(legend.position = "bottom")
    #ggtitle("Observed vs predicted Hb-values")
  return(scatter_plot)
}
```

```{r, include=FALSE}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")

@@

```{r}
g1 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, NULL)
g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.1)
if (save_figs) {
  ggsave(filename="classification-scatter-plot-hb-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g1, dpi=600, units="mm", width=180)
  ggsave(filename="classification-scatter-plot-probability-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g2, dpi=600, units="mm", width=180)
}
g1
g2
```

```{r, eval=FALSE, include=FALSE}
jarkko_test <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
